problem years

### 2024-06-18 src/problem_years.Rmd
corrected OAs

```{r}

@@

### Problem postcodes

@@

### Correct 1996-2010

@@

df9610$n[is.na(df9610$n)] <- predictions[is.na(df9610$n)]
df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)

df9610 <- as_tibble(df9610) %>% 
  mutate(year = as.double(as.character(year))) %>% 
  left_join(lookup, by = "pc", suffix =c("","")) %>%
  group_by(year, ladcd) %>%
  summarise(n = n()) %>%            
  ungroup()
```

@@

```{r}
data.path2011_2012 <- paste0(path, "/data/temp/domain_pc_year1112.csv")
#Created by domain1112.Rmd, which is based on domain.R and uses domain instead of host.
#This is what we use for the hyperlinks paper as per George's script

@@

## Firms

@@

## spatial data

@@

## lisa maps

```{r}

path.out <- paste0(path, "/outputs/lisa/corrected/lisa_level_pc", n, "_la") # for unique PC
#path.out <- paste0(path, "/outputs/lisa/corrected/lisa_level_pc", m, "_la")  # for multiple PC

@@

# Moran I

```{r}
path.out <- paste0(path, "/outputs/lisa/corrected/morani_la.csv")
morani %>% as_tibble() %>% 
  rename('morani'=V1,
                   p = V2,
                   year = V3) %>% 
  #kable() %>% 
  write_csv(path.out)
```

```{r}
path.images <- paste0(path, "/outputs/lisa/corrected/")

list.files(path = path.images, pattern = "^lisa_level_pc1_la", full.names = T) %>% 
  map(image_read) %>% # reads each path file
  image_join() %>% # joins image
  image_animate(delay=50) %>% # animates, can opt for number of loops
  image_write("corrected_lisa.gif", format = "png") # write to current dir

```

```{r}
# centers the local Moran's around the mean
lmoran$Ii <- lmoran$Ii - mean(lmoran$Ii, na.rm = TRUE) 
lmoran$lag.n<-  lag.listw(listw,lmoran$n, NAOK = TRUE)

# centers the variable of interest around its mean

lmoran$ns <- lmoran$n - mean(lmoran$n, na.rm = TRUE) 
lmoran$lag.n <- lmoran$lag.n - mean(lmoran$lag.n, na.rm = TRUE)

signif <- 0.05
#lmoran

lmoran <- lmoran%>% 
  mutate(quadrant= ifelse(ns>0 & lag.n > 0, 1, 0)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.n < 0, 2, quadrant)) %>% 
  mutate(quadrant= ifelse(ns<0 & lag.n > 0, 3, quadrant)) %>% 
  mutate(quadrant= ifelse(ns>0 & lag.n < 0, 4, quadrant)) %>%   
  #mutate(quadrant= ifelse(lmoran$`Pr(z > 0)` > signif, 0, quadrant)) #%>% 
  mutate(quadrant= ifelse(lmoran$`Pr.z....E.Ii..` > signif, 0, quadrant)) #%>%

#   mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st > 0, 1, 0)) %>% 
# mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st < 0, 2, quadrant2)) %>% 
#   mutate(quadrant2= ifelse(ANC_st<0 & lagANC_st > 0, 3, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(ANC_st>0 & lagANC_st < 0, 4, quadrant2)) %>% 
#  mutate(quadrant2= ifelse(lmoran$LISA_PANC > signif, 0, quadrant2))

mun_merge_new<- merge(test, lmoran, by="LAD21CD")
mun_merge_new<- test %>% 
  left_join(lmoran, by="LAD21CD")

# R p value map
breaks = c(0, 1, 2, 3, 4, 5)

cities <- maps::world.cities %>% 
  filter(country.etc=="UK") %>% 
  arrange(pop) %>% tail(10) 
cities.sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)

map <- tm_shape(mun_merge_new) + tm_fill(col = "quadrant", breaks = breaks, 
                                         palette = c("white","red","blue",
                                                     rgb(0,0,1,alpha=0.4),
                                                     rgb(1,0,0,alpha=0.4)), 
                                         labels = c("Not significant",
                                                    "High-High",
                                                    "Low-Low",
                                                    "Low-High",
                                                    "High-Low"), 
                                         title="LISA, websites per firm in a LA") +
  tm_legend(text.size = 1) +
  #tm_scale_bar(position = c("LEFT", "BOTTOM"),text.size = 1.0)+
  #tm_compass(type = "8star",   position = c("RIGHT", "BOTTOM"),      show.labels = 2,   text.size = 0.5)+
  #tm_borders(alpha=.1) +
  tm_borders(lwd = 0) +
  tm_shape(la, simplify = .2) + tm_borders(alpha=1, lwd = .7) +
  tm_shape(cities.sf) + tm_dots() + tm_text("name", size = .5) +
  tm_layout( frame = FALSE,  title = t) #"LISA with the R p-values ")
tmap_save(map, filename = paste(path.out, t, ".png", sep = ""))
}

#ggplot map solution.
# mun_merge_new %>% st_as_sf(crs = 27700) %>% #st_transform(4326) %>%
#   ggplot()+
#   geom_sf(aes(fill = as.factor(quadrant)))
# ggsave("test_ggplot.png", path = path)

```

>>>>

```{r, eval=FALSE}

@@

## Problem postcodes

@@

## Correct 1996-2010

@@

df9610$n[is.na(df9610$n)] <- predictions[is.na(df9610$n)]
df9610 %>% filter(pc == "M28 2SL") %>% distinct(year)
```

## LA

@@

```{r}
df9610 <- as_tibble(df9610) %>% 
  mutate(year = as.double(as.character(year))) %>% 
  left_join(lookup, by = "pc", suffix =c("","")) %>%
  group_by(year, ladcd) %>%
  summarise(n = n()) %>%            
  ungroup()

@@

### Firms

@@

### spatial data

@@

### lisa maps

```{r}
path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", n, "_la") # for unique PC
#path.out <- paste0(path, "/outputs/lisa/corrected/LA/lisa_level_pc", m, "_la")  # for multiple PC

@@

### Moran I

```{r}
path.out <- paste0(path, "/outputs/lisa/corrected/LA/morani_la.csv")
morani %>% as_tibble() %>% 
  rename('morani'=V1,
                   p = V2,
                   year = V3) %>% 
  #kable() %>% 
  write_csv(path.out)
```

## OA

```{r}
df9610 <- as_tibble(df9610) %>% 
  mutate(year = as.double(as.character(year))) %>% 
  left_join(lookup, by = "pc", suffix =c("","")) %>%
  group_by(year, oa11cd) %>%
  summarise(n = n()) %>%            
  ungroup()

df1112 <- read_csv(data.path2011_2012) %>% 
  filter(
    #V1.domain < m,             # for multiple pc
    V1.domain == n,             # for n == 1
    year > 1995 & year <2013) %>%   
  left_join(lookup, by = "pc", suffix =c("","")) %>% 
  group_by(year, oa11cd) %>%
  summarise(n = n()) %>%             # for n == 1  
  #summarise(n = sum(V1.domain)) %>% # for multiple pc  
  ungroup()

df <- bind_rows(df9610, df1112) %>% 
  filter(!is.na(oa11cd)) %>% 
  complete(oa11cd, year) %>% 
  replace(is.na(.), 0)

### spatial data

```{r}
# get OA for England and Wales
path.geo <- paste0(path, "/data/raw/Output_Areas__December_2011__Boundaries_EW_BGC.geojson")
oa.ew <- st_read(path.geo)
# source: https://geoportal.statistics.gov.uk/

# spatial transformations
oa.ew <- st_transform(oa.ew, 4326)  # EPSG code for WGS84

# keep in the data slot only the ONS Output Area id, renaming it as 'id'
oa.ew <- oa.ew %>% 
  dplyr::select(OA11CD, Shape__Area) %>% 
  rename(id = OA11CD,
         area = Shape__Area)

path.geo.sc <- paste0(path, "/data/raw/output-area-2011-mhw")
oa.sc <- st_read(dsn=path.geo.sc, layer = "OutputArea2011_MHW")
# source: https://www.nrscotland.gov.uk/statistics-and-data/geography/our-products/census-datasets/2011-census/2011-boundaries

# spatial transformations
oa.sc <- st_transform(oa.sc, 4326)  # EPSG code for WGS84

# Scotland (follows same steps as EW, see notes above)
oa.sc <- oa.sc %>% 
  dplyr::select(code, SHAPE_1_Ar) %>% 
  rename(id = code,
         area = SHAPE_1_Ar)

# NI
path.geo.ni <- paste0(path, "/data/raw/ni_small_area/")
oa.ni <- st_read(dsn = path.geo.ni, layer = "SA2011")
oa.ni <- st_transform(oa.ni, 4326)  # EPSG code for WGS84
oa.ni <- oa.ni %>% dplyr::select(SA2011, Hectares) %>% 
  rename(id= SA2011,
         area = Hectares)

# Build OA for UK
oa.uk <- rbind(oa.ew, oa.sc, oa.ni)
rm(oa.ew, oa.sc, oa.ni)

# get UK
path.uk <- "https://services1.arcgis.com/ESMARspQHYMw9BZ9/arcgis/rest/services/Countries_December_2023_Boundaries_UK_BUC/FeatureServer/0/query?outFields=*&where=1%3D1&f=geojson"
uk <- st_read(path.uk)
# spatial transformations
uk <- st_transform(uk, 4326)  # EPSG code for WGS84
```

### lisa maps

```{r}
path.out <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc", n, "_oa") # for unique PC
#path.out <- paste0(path, "/outputs/lisa/corrected/OA/lisa_level_pc", m, "_oa")  # for multiple PC

#spatial weight matrix
sf_use_s2(FALSE)
nb <- poly2nb(oa.uk)
listw <- nb2listw(nb, style = "W", zero.policy = T)

for (t in 1996:2012){
  test <- oa.uk
  test <- test %>% left_join(df[df$year==t,], by = c("id"="oa11cd")) %>% 
    mutate(n = ifelse(is.na(n), 0, n),           # 0 for NAs and the same for year
           year = ifelse(is.na(year), t, year))

  #LISA
  lmoran <- localmoran(test$n, listw,  adjust.x=TRUE, zero.policy = T, na.action = na.exclude)

  quadr <- attr(lmoran, "quadr")
  #quadr[lmoran[,5] > 0.05 | is.na(lmoran[,5]), ] <- NA
  quadr[p.adjust(lmoran[,5], method = "fdr") > 0.05 | is.na(lmoran[,5]), ] <- NA

  cities <- maps::world.cities %>%
    filter(country.etc=="UK") %>% 
    arrange(pop) %>% tail(10) 
  tif_sf <- st_as_sf(cities, coords = c("long", "lat"), crs = 4326)

  ggplot(data = oa.uk, aes(fill = quadr$median)) +
    geom_sf(data = uk, fill = "white", linewidth = .5) +
    geom_sf(linewidth = 0) +
    geom_sf(data = tif_sf, fill = "black", size = .3) +
    #geom_sf_text(data = tif_sf, aes(label = name, fill = "black"), size = 4) +
    scale_fill_manual(values = palette_new, na.value="white", drop = F) +
    theme_minimal() +
    theme(panel.grid.major = element_line(colour = "transparent"),
          axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          legend.title=element_blank()) +
          #legend.position="none") +
    guides(fill = guide_legend(reverse = TRUE)) +
    labs(title = t,
        subtitle = "LISA, websites per OA")

  filename = paste(path.out, t, ".png", sep = "")
  ggsave(filename, bg='#ffffff')
}
```

### Moran I

```{r}
path.out <- paste0(path, "/outputs/lisa/corrected/OA/morani_oa.csv")
morani %>% as_tibble() %>% 
  rename('morani'=V1,
                   p = V2,
                   year = V3) %>% 
  kable() %>% 
  write_csv(path.out)
```
